% Chapter 1

\Chapter{The University Side}{And internal actors} % Main chapter title

\label{Chapter2} % For referencing the chapter elsewhere, use \ref{Chapter1} 


\section{University as institution}
The first university has been founded in the 1088, but the \enquote{scientific} approach and role we usually acknowledge them for has come only in the 17th century. The \enquote{open science}, that has characterized their activity until now, is a \enquote{widespread system of exchange based on the value of scientific priority and prestige} \citep{Murray2005}. It relies on a series of norms that facilitate full disclosure and diffusion of research results, where economic incentives push toward a cumulative knowledge production, i.e. public research funding that recognize and reward the priority of the research outcome.

Lately in the previous century, a new system has arisen oppositely to the open science: the \enquote{patent system}, also called the \enquote{economy of inventions}. In this regime, property rights on scientific discoveries ensure the potential stream of financial rewards, based on the ability to exclude others form the appropriation of newly created value and the reward for commercial exploitation in exchange for the disclosure. 

Clearly there is a conflict between these two models, enforced by the increasing focus of universities and public policies on commercialization activities, further challenging the culture of open science. In fact, academic scientists' fundamental devotion to the open science paradigm created a third alternative to these systems, a hybrid economy with mixed elements from both parties. In this perspective, both firms and academic inventors use patents to protect and exchange new knowledge – even if it is believed to be \enquote{less efficient than open science} \citep{Geuna2009}. Similarly, \citet{OwenSmith2001} found a convergence toward a hybrid system of scientific and technological success.

An example could be the phenomenon of \enquote{patent-paper pairs}: starting from the same knowledge, a patent will be firstly applied for, then a paper will be publisher after the delay required by the patent process \citep{Murray2005}. In this model, the needs of both parties will be satisfied: the publication, for academic purpose, and the IP protection, for commercialization. In fact, even scientists themselves are interest in protect their ideas, as patents have become important bargain chip - a currency in the knowledge economy.

A different terminology is used by \citet{Stern2004} who refer to a \enquote{science approach} quite similar to the open science approach. He defined it also (and most interesting) through the peculiar reward system based on the scientific priority, which drives researchers to publicize their findings as quickly as possible, without retaining any right on the intellectual property. 

There is a widespread debate on which paradigm should be adopted by publicly funded institutions like universities, especially when considering the type of knowledge involved in each system: basic and applied research. Many organizations tried to mix these elements, but the tension among them lead to extensive differences in objectives and administration, including major difficulties in managing the conflicting goals of curiosity-driven research and commercialization activities, left alone the various shadows between these extreme points \citep{Rasmussen2006}.

A first argument could be the one of \citet{Geuna2009}, in favor of the new commercialization model, who stated that the mission of creating and disseminating \enquote{knowledge for its own sake} can drive university scientists away from real-world, practical problems that need to be solved. \citet{Rosenberg1994} and \citet{Nelson1998} also observed that the large part of disciplines that take part in the academic curricula were developed specifically to meet requirements of firms. \citet{Balconi2006} noted that universities must be \enquote{intimately familiar} with industrial technology, in order to identify and perform useful research. Moreover, the technology transfer is a two way knowledge flow, where academics can benefit too.

A more moderate approach has been taken from \citet{Beath2000} who reported that there is a trade of in the effort an academic scientist can devolves to the creation and dissemination of knowledge and its commercialization, starting from a limited amount of time and energy. Therefore, he seems to justify these latter activities only if they have an actual financial impact on the university budget. He arguments that due to its public good nature \citep{Muscio2013}, university knowledge should be treat carefully: he refers to the case of the pharmaceutical industry, even if this relates more with the question of patent legitimacy in \enquote{greater good} matters.

\citet{Siegel2003a} instead defend the open science system. Firstly, he reports interviews where the commercialization mission was seen as inconsistent, with the traditional public domain philosophy that should be endorsed by public institutions. Then, he refers to a trade-off stated in \citet{Nelson2001} where \enquote{the shift away from open science might slow down technological diffusion}. Also \citet{Stern2004} states that the priority system embedded into the open science approach seems relatively efficient, by discouraging shirking and encourages maximal knowledge diffusion.

\citet{Rasmussen2006} report various changes in the government control and administration of universities, generally aiming towards a greater autonomy and competitiveness through a performance-based funding. However, even if the change itself has no positive or negative connotation, he seems to agree with others, arguing that academic freedom and basicness of university research may be treated by commercial activities. Others expose their concern about the shorter time horizon imposed by a more applied research, various tensions and conflict of interests. 

\citet{Murray2005} takes a case study as exemplification of the actual attitude and reaction of academic and industrial scientists to the new patent system: the oncomouse case. Briefly, a company funded and patented a research that eventually led to one of the most recognized innovation in its field. However, the firm imposes strong limits in the acquisition and usage of oncomouses, the informal exchange of relative knowledge, and the appropriability of \enquote{reach-through} discoveries, leading to \enquote{a widespread infringement} from scientists. 

Similarly, \citet{Stern2004} empirically investigates this scientists' willingness to perform basic or applied research within these different systems, to get a better understanding of their true preferences. He studied the relationship between wages and the scientific orientation of R\&D organizations, founding a significant and positive correlation between the closeness of the employer organization's science approach and the premium wage they must pay.

Elsewhere, other authors criticize more directly the way universities are currently interpreting the phenomenon. \citet{Siegel2003a} consider the university management of IP rights too aggressive, and report a conflict between the overall willingness to commercialize and tenure and promotion policies, that promote scientific publications and grants. \citet{Murray2005} instead observe that an extensive protection of intellectual property require the negotiation of every exchange relationship, slowing down the transfer process.

Historically, the exploitation of research results has not been the first concern for most of academic institutions. As reported by \citet{Balconi2006}, in the previous century the technology brought to the market was merely an application of previously and independently developed scientific knowledge. However, since the early 1980s, universities have seen the emergence and consolidation of a \enquote{third mission}, alongside the research and dissemination objectives. This additional aim requires universities to bring research results into the market, not only for fund raising purposes (even if it is usually pointed as the main reason) but also for economic development, technical advancement and wellbeing. 

This orientation has led to the development of the concept of \enquote{entrepreneurial university}. This term was coined by \citet{Etzkowitz1998} to describe universities that have proven themselves critical to the economic development of their region. This idea is based on the academic entrepreneurship, defined in \citet{Louis1989} as \enquote{the attempt to increase individual or institutional profit, influence, or prestige through the development and marketing of research ideas or research based products}.

In this new economy setting, this knowledge-based entrepreneurship has become one of the driving force of economic growth: entrepreneurial universities act as a knowledge producer and disseminator, but also as a natural incubator where scientists and entrepreneur can explore, evaluate and exploit ideas, through the provision of an \enquote{adequate atmosphere} and a network of relationships necessary to the entrepreneurial activity \citep{Guerrero2014}.

To better understand this approach, it can be compared to the \enquote{research university}, whose main purposes are to (1) conduct research and (2) train graduate students to perform research. In fact, these universities are considered more effective in the knowledge transfer, relatively to federal laboratories \citep{Rogers2001}, but they still miss the acknowledgement of the growing role of knowledge and technology advancement in this era, jeopardizing the opportunity to become fundamental actors in the economic development of their regions \citep{OShea2004}.

Nowadays, the importance of entrepreneurial universities can be evaluated through the extent of actions taken by governments and other institutions, that are actively pushing universities to directly engage in technology transfer activities, i.e. contract and collaborative research. Contemporary, these can be seen as a recognition of the great endowment of knowledge-based assets owned by universities, that is seen as economically underexploited \citep{Tijssen2006}. 

\citet{Balconi2006} seen this recognition also in the trend of putting \enquote{institution of science} in charge for riskier explorations: his though is that the \enquote{ultimate sense} of these institutions is to set the foundations for new developments. It is implicit that the necessary condition for this is that universities must possess the necessary knowledge and a climate.

But what changed overtime in a more operative perspective? Early in the days, when the third mission was not so well established, the technology transfer heavily relied on personal relationships of academics, and there was only a few institutions that has units or offices to support these activities, usually with few non-specialized employees. New technologies for the market was developed prevalently in large companies' laboratories, while small and medium enterprises have other choices, i.e. cooperative research centers. What \citet{Geuna2009}recognize now, instead, is the large trend to professionalize the technology transfer and the third mission of universities, whereas industrial R\&D is getting more and more directed toward commercial ends \citep{Fritsch2007}.

But what he acknowledges as source of this discontinuity is a change in the context, especially in firms involvement. Specifically, the use of knowledge as a competitive advantage; the increasing demand for skilled employees; the availability of higher educational levels; the usage of universities as policy tools for economic development by governments; greater budget constraints and diminishing incomes from governments. Another perspective is that in the \enquote{scientificaton} of the industrial production process, universities and public research organizations represent the main knowledge producers, fundamentals to the economic development \citep{Balderi2007}. Lastly, already \citet{Thursby2002}found evidences to further support the importance of changes in faculty orientation toward commercialization and collaboration, in addition to a shift in industry R\&D.

\citet{Baldini2007} reinforce this view stating some non-obvious factors that has driven to a greater involvement of universities. First, innovative technologies reduced the need for concentrating research activities in large facilities that can afford certain costly machinery, allowing smaller centers to be competitive in performance and effectiveness. Secondly, the policy shift in allocation of intellectual property rights act as a motivation for researchers to get more involved. Third, he observes that these factors, unitedly to the declining amount of public funds, lead to an increasing competition for any source of funding. In this perspective, this process has become of \enquote{vital importance} for universities \citep{Muscio2008}; however, its use as funding source is a secondary objective \citep{Jensen1998}.

A relevant non-financial objective, for example, can be taken from \citet{Tijssen2006}: the willingness to achieve and maintain a leadership in scientifically relevant fields. \citet{Leitch2005}, instead, cite as a driver the desire to develop, retain and acquire skilled graduates. \citet{Baldini2006} report the importance of technology transfer to reinforce the university reputation. 

In this setting, \citet{Wong2010} investigates the impact of technological commercialization on publication activities: he found a significant and positive correlation between patent applications and publication output, suggesting a mutually reinforcing relationship. \citet{Lee2000} listed other potential individual motivations: to test researches, gain insight and knowledge on practical problems, for teaching purpose, students internships and placement, to look for business opportunities. 

Closing the stakeholder perspective, governments need universities to accomplish various tasks, especially referring to the regional innovation system \citep{Fritsch2007,Balderi2007}: to generate, accumulate and transfer knowledge to local actors, i.e. through R\&D cooperation and innovation-related services; develop the local labor force; strength the absorptive capacity of the region and firms; incubate new technology-based firms. \citet{AzagraCaro2010} directly cited as advantages, among these, the develop of new technical instruments and methodologies, the enhancement of problem-solving capabilities in the local context, the formation and stimulation of networking activities and social knowledge. Another reported advantage is the spinoff formation, a highly prioritized area for governments.

In fact, universities have a \enquote{special} attention for matter of national institutions: \citet{Rasmussen2006} observes that universities \enquote{seem quite eager to satisfy public expectations, which in turn generates goodwill from research councils and ministries}, referring to the issue of decreasing funding. This expose the need to show actual, short term results both in traditional missions (publications and students) and the direct contribution to the national economy, as described later.

These forces have led to a true change in the attitude of universities; \citet{Tijssen2006} describe this shift toward 3 different phases of development. The first is the \enquote{application oriented - science driven} approach, that coincides with the rising awareness of possible link between universities and companies for the resolution of industry-related problems. Later comes the \enquote{product oriented - utility driven} approach, that consists in the emergence of activities such as prototyping, development services etc. specifically aimed to the discover and exploitation of opportunities. Third and last is the \enquote{business oriented - market driven}, where IP rights are secured, relations with firms are established and first products are sold.

In this perspective, great distance has been put between the traditional Ivory Tower image - usually associated with the 80s and 90s typical university - and the actual industry-friendly university involved in the solution of real-world problems \citep{Baldini2006}. However, must be mentioned \citet{Rasmussen2006}, whose empirical findings indicate \enquote{soft emphasis on commercialization}.

What this historic perspective cannot illustrate are the challenges that universities are facing in this new system; \citet{Rasmussen2006} divide them in three different groups First, to increase the extent of commercialization activities and outcomes, to produce a greater impact on the innovation system and gain a better financial perform. Secondly, to visualize the contribution to the economic development, both at a local and a national level, to justify the university's funding requirements with their economic role and social impact. Lastly, to overcome the various difficulties in the management of the new fundamental activities: teaching, the exogenous (as driven by curiosity), and endogenous (market-driven) research \citep{Debackere2005}. 

Summing up, universities must cope with multiple, greatly different mission required by stakeholders with different purposes. What is needed, to overcome these challenges, is ambidexterity: universities nowadays should not draw a clear line between exploration and exploitation activities. They should manage their relative tension, integrate them and take advantage of every potential synergy.

\citet{Chang2016} develop this perspective proposing two levelss, or types, of ambidexterity. The first is the top-down, \enquote{structural} ambidexterity, that uses organizational tools to promote synergy exploitation, such as task partitioning, separation of tasks and units, an appropriate leadership. The second is a bottom-up, \enquote{contextual} approach, that aims to develop an appropriate organizational climate to encourage individuals in directly managing the conflict.

In refer to this ambidexterity capability, many authors assume an historical perspective to identify and evaluate the most common constrains for its development: the very scientific method of universities. These institutions have developed over time a series of values, beliefs, norms of conduct and organizational practices devoted to the open science paradigm that can actually impede an effective push toward commercialization activities. 

\citet{Argyres1998} refer to this problem as the \enquote{social-contractual commitment of universities to the intellectual commons}, which constraint their ability to exploit economic opportunities. \citet{Muscio2013} instead observes the tension between two extreme points of a continuum: the ideal norms of open science and the Mertonian ethos, which have long driven activities and evolution of the academia, and the modern commercial approach, filled with business values.

In this perspective, every university has its peculiar historical path. In fact, many authors found these as one of the most relevant factor in their comparative analysis of technology transfer performance. As an example, \citet{OShea2005} identify the \enquote{uniqueness of historical conditions} as the basis for a sustained competitive advantage, when evaluating spinoff performances. 

Many empirical researches support this point, confirming the importance of the previous performance in technology transfer in promoting a higher engagement. Another relevant research on this topic is the one of \citet{Baldini2006}, who linked the behavior of potential inventors and entrepreneurs to the historical performance, observing that processes as patenting can be \enquote{cumulative}, due learning effects. Finally, previous competitive performance can reinforce the institutional (and governmental) attention to the commercialization activities, further pushing them.

However, has to be specific remembered that there are also empirical evidence to discredit this correlation, at least in particular perspectives, like the results showed by \citet{Thursby2002} who find evidences for a little effect of past (licensing) performance on the propensity of the university administration to further promote (patenting) activities.

%----------------------------------------------------------------------------------------

\subsection{Performance and attitude}

Other factors make universities differ in the degree they engage with industry, and many authors have built formal models to uncover which of them are significant, and evaluate their impact. While the econometric results will be reported later in chapter 4, it is relevant to introduce now which university's characteristics can play a role in determining the technology transfer performance.

As a starting point, the most recognized factor that influence the university inclination to engage in such activities is its commercial orientation \citep{DEste2007}, or the \enquote{catch-all phrase entrepreneurial culture} in \citet{OwenSmith2001}. It is intended as the sum of its cultural legacy, its history and its past performance and to what extent these factors influence the nature of the academic research. In fact, the commercial orientation in its definition is so close to the technology transfer willingness and engagement, that it can be interpreted both as the dependent and independent variable. However, it is mainly recognized as an enabling factor that sets a fertile ground for commercial and business activities. 

Strictly related to the commercial orientation, are a series of factors that can be referred as the supportiveness of the university environment. Examples are various policies for academic entrepreneurship: leave of absence, access to laboratories and other infrastructures, entrepreneurial programs, senior member championing for entrepreneurial projects, on-campus incubators, investments in startup, earlier and easier access to venture capital \citep{Baldini2007}. \citet{Guerrero2014} uses similar variables in his model, such as entrepreneurship education programs, the attitudes toward entrepreneurship of key actors in the hierarchy, a strong top-down oriented leadership, the availability of an entrepreneurial model.

Similar, but loosely coupled factors are variables linked to the perception of the university orientation, rather than influencing the context themselves as the environment supportiveness. An example is the research by \citet{DEste2007}, on whether the founding mission specifically includes the university support to regional development; in fact, he found a significant and positive relationship with commercial engagement.

The literature has considered other university-level factors that refer to the actual, processual ability to support the technology transfer process, rather than its ability to motivates academics and grow a supportive environment. Examples are the size of the research staff, the quality of the research production, a sizeable budget for entrepreneurial related activities \citep{Colombo2010}, the university own financial, technological and social capital, the institution status and prestige \citep{Guerrero2014}.

Among these, many research aimed to uncover the influence of organizational structures and governance on the engagement, which have been found to have a significant impact on the ability to manage internal conflict between missions, thus on the overall performance. \citet{Siegel2003a} found in its performance analysis that there is deviation from the best performance that cannot be completely explained by environmental and institutional factors, which should be attributed to organizational factors. Specifically, two level of organizational configurations has been studied.

The first is the internal organization of academics and research units. A specific structure that has been found improving the commercialization performance is the so-called research division \citep{VanLooy2004}: research personnel can decide to organize themselves and their research, alongside their contract research and other exploitation activities, in autonomous groups called research divisions, usually professionalized on a single topic or a specific industry. Performance of academics involved in such divisions has been found to be superior, and the gap widens over time.

The latter configuration refer to the presence, position and role of a dedicated office to technology transfer, henceforth TTO (Technology Transfer Office), which effectiveness has been found to depend on organizational practices \citep{Siegel2003a}. Generically, the institution of new, dedicated units and the importance vested in them are found to be relevant; other internal and external units or agencies that can help are the Technology Licensing Offices, Industry Liaison Office, incubators, science parks, joint ventures, spinoffs \citep{Tijssen2006}. More information on the topic will be provided later, but it is essential to understand the role of the university administration.

A related topic that has gain an increasing attention refer to policies regarding the evaluation of academic research personnel. Specifically, the reward systems, both monetary and not \citep{Guerrero2014}, criteria for career advancement, tenure policies, royalties allocation and alike. As a general rule, incentive policies should be promoted to raise awareness around the benefit in protecting and exploiting intellectual properties, to ease the perception of the patent process and increase the esteem scientists have of TTO's competence. This either will be discussed later, but it should be recognized from the beginning that a major issue is whether researchers have sufficient incentives to disclose \citep{Debackere2005}.

Lastly, local and national contexts universities live in plays a relevant role too. Examples are the relationship between annual public expenditure in university R\&D \citep{OShea2005}, the national economic growth \citep{Siegel2003a}, national culture and academic socialization \citep{Bercovitz2006}, the extent of R\&D activities of local firms \citep{Siegel2003a}, and the strength of ties between universities and industries developed in past projects (Powers, 2005). 

\section{Academic scientists}

The development of new innovative knowledge and cutting-edge technology critically depend on the quality of human resources involved. Star scientists and academics form top tier universities can reach discoveries faster, through an easier access to the personnel the needed specializations rather than the necessary technical resources \citep{OShea2005}. Apart from the ability to reach innovations, academic inventors can also play a positive role in commercialization activities, i.e. providing surrounding knowledge or leading themselves the process through a spinoff.

However, scientists differ in their willingness to get involved into technology transfer. Many researches have been performed to identify which factors can foster academics involvement, and what are their basic characteristics and traits. The basic framework for an appropriate categorization and understanding of these differences come from \citet{Stokes1997}. He divided academics in three different categories, according their interests and orientations. 

The first is what he called the Edison scientist: an academic who is interest only in purely applied research, oriented to the development of new product to meet people's needs. At the other end, there is the Bohr scientist that was intended to be a pure, basic scientist who most strictly adhere to the open science approach. Between these extremes, Strokes identified the Pasteur scientists as the most relevant for the technology transfer process, who presents as a main trait the desire to advance scientific knowledge, but only in fields or applications that can have real-world, useful applications. Ideally, this academic employee had authored many high-quality scientific papers, as well as applied for many patents; he should be a recognized expert in its fields, and both inclined to develop inventions and to gain a strong reputation in the scientific community through research activities \citep{Baba2009}.

Later studies use a dual categorization, useful even if simpler, in which scientific occupations have been divided into two distinct types \citep{Beath2000}. On one hand university scientists perform fundamental and basic research, motivated by the desire and the reward of establishing their priority on the discover. On the other hand, private sector researchers which has no opportunity for fundamental research; they are financially motivated in keeping up with the scientific literature and transfer it into marketable products. 

Both these perspectives match what previously was referred to as the difference between the open science approach and the IP protection system: Bohr scientists, \enquote{university} scientists \citep{Beath2000}, and partially Pasteur scientists share a common orientation toward the open science; Edison and again Pasteur scientists tend to commercialization activities. Therefore the peculiar and extremely relevant role of the Pasteur scientist, who is in the position to bridge interests, knowledge and activities.

Specific to the dual perspective is the fundamental study of \citet{Stern2004}, who investigate the impact of the organizational orientation on scientists' wages, comparing public research organizations and private firms. He starts form two different assumptions: researchers may have an intrinsic preference toward the open science approach, while firms may have economic interests and benefit for participating to the same paradigm. He based his research on the tightness of control over IP, non-disclosure clause and similar variables.

His empirical results show that, once controlled for the relative ability of researchers, there is an actual premium wage payed by \enquote{closed} firms devoted to the protection of ideas, and that it is considerable in its extent, evaluated in around 27\% of difference in wages. In fact, scientists pay a compensating differential for participating in the open science system. 

What seems to motivate and drive the academics in this perspective is their unique sensitivity to reputational awards: an increase in reputation among colleagues, in an open science system, open the path to better research projects, teams and memberships, ultimately leading to employments in more prestigious institutions and access to relevant resources. This effect, as in other examples of the Matthew effect, seems to increase over time.

\citet{OwenSmith2001} recognize in their work this difference in a slightly different perspective, addressing two type of scientists from different areas: life science and physical scientists. These labels might correspond to basic and private researchers, respectively. The authors find that both types of scientists express concerns for various constrains due to commercial exploitations, and they still may engage but for very different objectives. The private scientist (including commercial-oriented academic) seek IP protection mainly to pursue commercialization activities while maintaining the original, unexploited value of the new technology. The latter category, representative of a purer researcher, may also seek IP protection, but with the very different intent of \enquote{shielding the environment of his lab from encroachment by commercial interests} and ultimately as a leverage to attract investments.

The his open science approach, still the most diffused today, shapes various factors that influence the academic willingness to engage in commercial activities. The main group of variables refer to career concerns: the system heavily relies on reputational awards, based on priority, novelty and quality of the research, rewarding scientists with a greater access to resources and better personal networks. This drive scientists to construct their research agenda freely form external influences that might carry unwanted constrains in time and effort. This is partially due to the \enquote{ambiguous relationship of researchers to money} \citep{OShea2004}, that refer to the disinterest nature of university research and thus the seek for public funding that enable academics and teams to work completely unrelated and free from industry-related constrains. 

A recognized factor related to this perspective is the unwillingness to delay publication of many researchers \citep{OShea2004, Thursby2002, Baldini2007}. The main driver is the \enquote{publish or perish} rule, that require academics to continuously publish in order to gain, and not lose, reputation. These delays may arise form: bureaucratic issues, i.e. time and effort required in disclosing the invention to the TTO; commercial issues, i.e. establishing a relation and to reach an agreement with firms; legal issues, i.e. the time required by the patenting process. Moreover, agreements with firms may require secrecy or a fixed-length publishing delay to allow the industrial counterpart to gain a first mover advantages; these requirements clearly threat publication activities, but also networking activities. 

These factors ultimately drive faculty member to not disclose internally their invention, thus not engaging technology transfer, because of \enquote{their believe that commercial activity is not appropriate for academic scientists} \citep{Bercovitz2006}, referring to the norms of open academic science.  In this perspective, the university administration acquires a relevant role: empirical findings suggest that a relevant part of the faculty resistance in disclosing is due to university-level policies focused on scholarly woks, while commercial activities are believed to be perceived as mere, non-relevant \enquote{services} \citep{Markman2005}. Examples are factors considered for tenure and promotion decision.

\citet{Muscio2013} express this duality of imperatives, the open science paradigm and the commercialization need, as a cognitive dissonance that academic scientists may directly experience when trying to reconcile these conflicting aims and their own research agenda. He noted that even if the cultural environment and the organizational context do matter, scientists' decision on which research to perform may largely depend on their personal propensity toward a specific approach.

These factors may keep scientists from commercialization activities, but there are actual benefits that should positive influence the willingness to engage. In fact, even the impact of commercial activities on publication has been largely discussed in the literature. A fundamental study on this topic is the one from \citet{Lee2005}, which showed that researchers who collaborate with firms are generally more productive in terms of publication, and the quality of journals not diminished. \citet{Lebeau2008} found empirical support that industry collaboration increases academics' citation indexes when normalized by field, while on a first sight the average relative impact factor could be diminished; oppositely to Lee and Bozeman, however, he found that journals published in have a lower impact factor, while articles receive higher citation counts.

Many authors agree on this view. \citet{Meyer2006} found that academic inventors tend to perform better in publications than pure academic scientists; \citet{Wong2010} found a positive relationship between research performances and patenting activities, used as a proxy of commercial activities. Elsewhere has been found a positive relationship between the quality of research, measured through publications and citations, and the extent of collaboration with industries \citep{AzagraCaro2010} or the patent productivity both at individual and university levels \citep{Baldini2007}.

How to reconcile these views? On one hand, academics are afraid of constrains that commercial activities may impose, and the negative impact that could have on their career; these worries may arise from poor information and a low awareness. On the other hand, industry collaborations do have a positive impact on research activities and academic careers, even when it implicates some limits.

\subsection{Individual characteristics}

As stated, the open science culture may have a negative effect on scientists' engagement with industry; however, many factors have been found to influence on their willingness. Apart from the institutional impact presented earlier, the literature analyzed many individual characteristics, while seeking for the determinants of researchers' entrepreneurship.

The first is a sort of cliché, which suggests that in order to have a good performance in commercialization activities, the academic should be \enquote{the entrepreneurial type}, identified as \enquote{who have always wanted to start companies and who use their university inventions as a way of achieving their entrepreneurial goals} \citep{Fini2009}. In fact, \citet{DEste2007} found that the most significant individual characteristic is the previous experience in collaborative research – a proxy for past behaviors – suggesting the presence of a Matthew Effect and thus the importance of the initial attitude toward industry.

Individual attributes usually associated with the academic entrepreneurship are \citep{OShea2004}: outgoing, extrovert personalities; a strong need for achievement; the desire for research independence; tenure and occupational skill level \citep{Roberts1991}; age and scientific experience \citep{Audretsch2000}; higher citation rates and publication performance \citep{Zucker2001}.

Elsewhere, academic life cycle models suggest that academics are more prone to technology transfer activities later in careers: in the early stages, academics work to increase their reputation and to build more extensive and valuable social networks; later, reached a satisfying level, they try to building legitimacy for their invention \citet{Fini2009}. However, empirical findings from \citet{DEste2007} suggests a higher probability of engagement for younger academics, possibly due to the minor devotion to the open science system and a different need for network development \citep{Bercovitz2006}.

Obviously, an extremely relevant factor is the environment researchers work in; as previously noted, a supportive university organization and culture is fundamental to improve the commercial involvement. As an example, \citet{OShea2004} found that the decision to start a spinoff is socially conditioned by the consensus that commercialization activities gain and the relative behavior of colleagues, as proxies for social norms, expectations, and the presence of prior academic entrepreneurs. 

\subsection{Academic motivations}

These characteristics describe what many empirical research have found as a typical academic scientist prone to technology transfer activities. But what motivates them? 

Many authors agree that the main motive for engaging with firms is not the entrepreneurial attitude, as might appear from the individual-level characteristics stated before. Instead, the main driver should be the expectation of a positive influence on their academic position and their research \citep{Fini2009, DEste2011}, according to the large adhesion to the open science approach.

Factors that directly influence the individual can be a gain in visibility, network development, prestige and reputation, recognition by peers \citep{Baldini2007, Fini2009, Rizzo2015}. In fact, the primary motivation for the academic scientists seems to be the recognition among the community and the reputation gain: collaborating with industry can help publishing in top-tier journals, conferences, federal research grants.

As largely recognized, technology transfer positively influence the research performance through the bi-directionality of information flow \citep{Geuna2009}, even if often neglected \citep{DEste2007}. Examples are the access to industry skills and facilities, opportunity for hypothesis testing, to verifying the applicability of the research, to keeping up with industry problems; new stimuli and new topics, etc. Generically, these reasons refer to the opportunity to interact with the external, real world \citep{Baldini2007}. 

Other motivations refer to the environment. Leadership in example can be effective as guidance, but also alter through its actions what is perceived as socially desired. The cohort effect \citep{Bercovitz2006} refer instead to the presence of a positive, previous experience of those in the same position, that act as a \enquote{case study} and affecting the willingness to engage. The same author found relevant a previous experience, especially a formal academic training (i.e. the PhD itself) in an institution which present a good performance in technology transfer. Similarly, \citet{Murray2004} and \citet{Link2007} demonstrate the importance of social capital in enhancing the probability of a successful cooperation with industry, therefore the importance of a supportive social network surrounding the scientist.

Different are the motivations that work at the individual level, but refer to the overall university. Cooperating with industry can bring more research funds, laboratory equipment, federal and European founds and grants, even attract star scientists to the institute \citep{OShea2004, Baldini2007, DEste2007}. However, the effectiveness of these motivations largely depends on the university reward system, i.e. tenure and promotion policies and royalty distribution formula; if not well managed, this system could lead to an institutional conflict between basic and applied research. 

This conflict is greater when disclosure is elicited and researchers are driven by exogenous motivations, rather than endogenous and indirect ones. Many author address this issue modelling the technology transfer process as a game between academics, TTOs, university administration and firms, and found a significant moral hazard problem for inventors \citep{Jensen1998}: their effort cannot be effectively monitored or enforces, which means that motivations for scientists should come from fine-tuned policies and the overall environment, rather than enforced by some internal regulation or by the hand of TTOs or administrations. 

In this perspective, compensation forms differ in extent and direction of their impact on researchers' motivations; expectably, the preferred form is the sponsored research, which among other channels allow scientists to continue work as they prefer, i.e. in-house with their own teams, on selected topics. Another suggested compensation is the equity share, that \citet{Jensen1998} found to be of less distortion on firms' and academics' decisions. Further attention should be placed in the form choice, especially when it comes to the various, unwanted effect of each channels and the conflict between them: an example form \citet{OShea2004} is that higher royalties allocated to academics will decrease the spinoff activity, by modifying opportunity costs, their ratio, thus their appetibility.

Along with compensation forms, it should be noted that even if a key objective of academics is the recognition within their community, i.e. through papers, presentations, conference and research grants, they are also motivated by personal financial reward and additional funding for their research \citep{Siegel2003a, Link2007, Fini2009}. \citet{DEste2007, DEste2011} agrees on the point, stating that fund rising performance can act as a signaling mechanism, positively affecting the reputation. \citet{Rizzo2015} adds two financial motives specifics to the spinoff path: funding for research, and tax avoidance. Greater financial incomes for the researcher or his unit may also lead to a greater independence in research lines.

In fact, \citet{DEste2011} found a positive attitude of researchers to financial ties with industry (74.5\% of interviewees), as long as funds are related to their research topic of choice and the open science paradigm is respected, i.e. when disclosure is agreed upfront and ideas are freely publicized. This study also confirms the priority of research-related motivations, thus the importance of a compatibility between financial and scientific rewards \citep{Link2007, Baldini2007}, and the need for a match between personal ambitions and the financial and business opportunity \citep{Tijssen2006}.

Finally, even if universities are seen as \enquote{professional bureaucracies whose members are relatively free to pursue activities that they believe are in the overall interests of the organization} \citep{DEste2011}, there is a clear need for pushing inventors to actively engage in the commercialization process \citep{Jensen1998}. Apart from the obvious intimate knowledge they can provide on the discovery and the surrounding scientific context, their passion in their work can make them assume the role of project champion, as well as helping TTOs in identifying and contacting companies, ultimately maintaining the relationship among the institution and the firm \citep{Markman2005}.

Summing up, also individual academics should demonstrate ambidexterity, as the attitude to simultaneously achieve publication and commercialization goals, starting from the ability to recognize exploitation opportunity from their research results \citep{Chang2016}. Individual motivation seems fundamental to accomplish both academic and entrepreneurial results, led by the comprehension of synergies and the mutuality of benefits for academics and firms. Similarly important seems the environment, which should balance emphasis and provide social support. 

\section{Technology Transfer Offices}

Intermediaries between suppliers and users of knowledge progressively emerged as central in bringing academic research to market \citep{Landry2013}. Universities and firms belong to different communities and react to different incentives, establishing a social and cognitive distance that require a specialized intermediator that can comprehend and effectively relate with both sides. The generalized acknowledgement and recognition of this problem led to a recent and substantial increase of investments, both from universities, firms and governments \citep{Muscio2010}.

The main source of their importance is the potential market failure involved in the transaction of newly developed technologies and knowledge: firstly, it is hard to identify which of the disclosed invention is actually marketable, due to the intrinsic innovativeness of the research output. The fundamental question, in this stage, is whether or not the idea has enough appeal for the market.

Secondly, if the invention match a potential market, there is no guaranty that the knowledge or the idea will grow in a functional prototype, due to high uncertainties embedded in the research and development process; even if this is available, it may not satisfy the industry requirement for a profitable production. In this case, the question to answer regard the feasibility of the project.

After these first, necessary leaps, the technology can be transferred to the industry. Two main issues may arise in this passage: to identify a potential, interested partner, and to establish a value for the transaction. Specific to the latter issue, there is the problem of estimating a new, unknown technology, which involve uncertainties and opportunisms and may lead to true market failures, i.e. great differences between social values and contractual prices, investments that fail to take places, let alone a failure in starting the transfer process.

Even if these issues arise from the generic problem of exchange and communication between research organizations and industry exponents, and even if a specialized organization can undertake any of these issues, the literature has mainly seen the resolution of the first three issues as \enquote{second order} topics, in researching the impact of intermediaries in latest problem: the focus is on their ability in, and the positive effect of, reconciling university and industry perspective, and to drive them to a mutually beneficial agreement; the implicit hypothesis is that the same factors that will allow a competitive performance on this topic, will be the same that allows the overcoming of all issues.

This specific interest of the literature arises from the recognition that the main and greater market failure come from the difficulties in, thus the imperfection of, estimating technology \citep{Hoppe2005}. This is mainly due to a problem of asymmetric information: firms usually cannot assess a priori the invention potential, while researchers have difficulties in identifying exploitation opportunities and evaluating their potential \citep{Debackere2005}. 

In fact, researchers have the best knowledge on the technology, while firms about the market; however both are required, to a successful and profitable transfer, but their match is constrained by limits typical of the agency theory. Similarly, \citet{Bercovitz2006} observes that differences in estimation may arise from subjective expectations on the knowledge value, while \citet{Hoppe2005} cited the uncertainty about the technology profitability. \citet{Debackere2005} further identifies the source of these constraint in the high-uncertain and non-codifiable nature of the scientific knowledge.

Within this perspective, intermediaries can enhance the overall system performance by reducing the asymmetric information issue: their professionalization, therefore their expertise in successfully locating and screening new ideas, can actively reduce the risks perceived by industries while providing indications to scientists on which idea to develop \citep{Debackere2005}. Moreover, intermediaries can balance the low bargaining power of individual scientists \citep{Bercovitz2006}, and also overcome the opportunism problem.

To uncover this last, specific benefit, some researchers have modelled the activity of intermediaries (mostly TTOs) as repeated games. An example is the one of \citet{Hoppe2005} who ultimately found that in a repeated licensing game, TTOs fully benefit only from high-quality inventions, therefore inducing them to push toward an equilibrium in which intermediaries sell only profitable inventions – if the disclosure frequency is high enough. In this setting, rewards for scientists are strongly suggested to be success-based, because fixed payments will not sustain the equilibrium. 

\citet{Macho-Stadler2007} comes at the same conclusion: in a repeated game, intermediaries have the incentive to \enquote{behave honeslty} to build a valuable reputation, while if the disclosing rate is insufficient or if the game is single and non-repeated, intermediaries may prefer to take advantage of the information asymmetry. Elsewhere, individual agents and small intermediaries may have the incentive to offer low quality inventions, whenever potential investors outnumber profitable inventions: scale is an issue \citep{Hoppe2005, Macho-Stadler2007}. 

However, there is a natural market force that push toward the concentration of intermediaries: first, a critical mass of inventions, thus the related research activity, is required to achieve a relative good performance, therefore the survival of the intermediary. Secondly, the economic performance of these actors is tightly linked to the ability, and the relative cost, of personnel; the larger the intermediary, the greater the opportunity to attract valued human resources and to economize on this key element.

Third, even if specialization is common in markets with intermediaries, the same force could lead to inefficiencies: in this peculiar case, and limited to the specialization by thematic, its low effectiveness in offering a balanced share of opportunities among intermediaries will eventually lead to the waste of high-quality invention, thus a market failure, and the underutilization of other intermediaries' resources, thus a market inefficiency \citep{Hoppe2005}.

Moreover, intermediaries with a larger invention pool may have incentive to further invest in experienced professional, more capable of locating new and profitable inventions. Apart from the obvious, immediate comparative advantage they could generate, their deployment may be necessary and their costs be more effective due to the size of invention pool \citep{Debackere2005}. In fact, economies on the costs of expertise generate one of the economic raison d'être of intermediation \citep{Hoppe2005}.

For specific forms of intermediaries, i.e. when they are internal to or participated by public research organizations, they are also required to manage the patent portfolio, requiring again additional personnel like patent attorneys. Empirical findings have demonstrate the importance of these specific professionals \citep{Siegel2003a}, in order to protect and market inventions and securing additional funds; moreover, the patent portfolio comes with management costs, which require a proper valorization activity to make it profitable \citep{Balderi2010}.

Lastly, intermediaries can also support the creation of spinoff, through their industrial linkages and by building a synergistic network between academics and venture capitalists, advisors and managers \citep{OShea2004}. Through these linkages, they have also the opportunity to identify business needs and to forward them to scientists; however, this activity will further increase the need for specialized personnel for additional marketing competencies \citep{Geuna2009,Muscio2008}. 

Summing up, intermediaries can overcome the main problem of universities: their bureaucratic organizational culture makes them inflexible in structuring deals \citep{Siegel2003a}, thus keeping them at a great cognitive distance from the market. In fact, intermediaries are becoming central, even if they may present some lack in effectiveness to most research institutions \citep{Geuna2009}, being them universities, research institutions or private research organization.

\subsection{Organizational forms}

Intermediaries may assume different organizational forms, due to different purposes, property settings, internal configuration and other factors. Examples are universities' TTOs, including TLO, ILO, and other nomenclatures; colleges; public research organizations, publicly founded agencies, knowledge intensive firms, professional associations, knowledge workers. Many researchers have investigated this issue, tackling it from various perspective, with the aim of establishing an ideal, effective form for every type of intermediaries. For the purpose of this work, apart from a general introduction, the focus will be centered on intermediaries strictly related to universities.

As a general categorization, an ideal starting point is the work of \citet{Yusuf2008} who divided intermediaries by purpose: general purpose, including universities, which aims to disseminate new knowledge; intermediaries specialized in helping various actors in the technology transfer process, as TTOs and alike; financial intermediaries, specialized in providing financial resources to projects, as venture capitalists and angel investors; institutional agencies, which provide incentives in order to promote the general economic development. Another useful classification comes from \citet{Landry2013}, which identify as \enquote{emblematic types} university TTOs, community college TTOs, public research organizations and non-profit organizations.

In fact, both examples have their roots in the same idea: different organizations need different services, therefore intermediaries must assume different forms based on specialized resources and capabilities. To make an useful example of these differences, \citet{Landry2013} turns to the knowledge value chain construct: the development of an idea can be divided in three non-linearly consequential stages, namely exploration, technical validation and exploitation. Each stage requires a different support from intermediaries, i.e. consulting on market needs, prototyping and patenting, commercialization activities etc. In fact, the author found significant differences in the level of engagement among institutions and intermediaries in the knowledge value chain stages: public research organization in exploration, community college TTOs on validation, non-profit organizations on exploitation, while university TTOs provide less customized solutions.

In the case of universities' dedicated offices, the previously cited Technology Transfer Office is in fact an oversimplification of the phenomena. Other real-world examples are the Knowledge Transfer Office, Industrial Liaison Office, Office of Technology Licensing, the University Technology Transfer Office and many other. However, regardless of their specific labels, these offices all perform variations of the same activity, with minor changes in focus and processes \citep{Brescia2016}; the only major difference is that some of these support prevalently the licensing process, while others are more general purpose. However, the main tendency is to \enquote{resize the importance of licensing income, to rather increase the effort in maximize industry-fund research} \citep{Balderi2010}. In this perspective, this difference may not last long. 

Instead, a significant difference among universities' TTOs is the organizational form, role and position they assume in different institutional context. The organizational form is the first factor that has been found by many to be a fundamental in determining the TTO performance; specifically in the case of a public entity like a university, in which academics' devotion is for the open science paradigm, an appropriate structure should allow the provision of both competitive mechanisms and adequate incentives to disclose. 

In fact, \citet{Bercovitz2001} propose that any transfer activity is \enquote{shaped by resources, relationships, autonomy and incentives of the TTO}, therefore the process outcomes may largely depend on organizational practices. This view is reinforced by \citet{Debackere2005}, which stated that different organizational arrangements may lead to different propensities of academics to commercial activities, especially for the \enquote{professional bureaucracy} previously cited. 

This traditional approach is based on a single, centralized office which employs a variable number of professionals, which can be organized by activities, projects, or none. In larger institutions, or where the attitude toward the market is greater, the office can grow enough to justify an internal divisional form, usually based on tasks similarities. However, empirical researches suggest that most universities employ only few employees with almost identical tasks. This traditional model can be seen as a divisional structure in which a single unit, the TTO, is in charge of the exploitation strategy. 

The second most common organizational form is the decentralized model, which is usually linked to a higher engagement in university-industry cooperation. Although there is no formal, shared definition for this model, it is mainly recognized as an ensemble of technology transfer officers located in the various research units, teams, faculty and alike. In this case, officers will be in charge for every aspect of the technology transfer, in respect to their research unit. This model refers to a matrix structure \citep{Debackere2005}, where the exploitation division became integrated within research groups, while responding for their activity to the central administration.

Both models have some limitations. On one hand, the centralized structure may induce the TTO to behave as a gatekeeper, to constitute a bureaucratic step toward the market instead of a supportive actor. Complex internal communication flows that come with a central structure, moreover, can slow down the process and make the university unresponsive to business and market needs \citep{Litan2008}. On the other hand, specialization and decentralization can ensure a higher responsiveness, positively influencing academics through proximity and avoiding possible conflict of interests \citep{Debackere2005}. However, this organizational solution compromise the exploitation of synergies among activities and topics, other than economies of scale embedded in a single office.

Anyway, an organizational form must be chosen. A comparative evaluation can be taken from \citet{Bercovitz2001}, who analyzed the potential impact of various Chandler's organizational forms on the office performance, both theoretically and on a case study basis. He found that the matrix form (corresponding to the decentralized form) provides a better coordination among internal actors, while ensuring the best incentives alignment. 

Later, a third possible model has been described by \citet{Brescia2016}: the semi-centralized model, as a compromise between the previous extremes. In this case, activities are distributed among a central TTO, typically in charge of intellectual property and spinoff activities, and decentralized officers employed for grants, collaborations and contracts. Theoretically, this configuration should allow the exploitation of advantages of both the previous models. 

It should be noted that the TTO may not be internal to the university. \citet{Fisher2002} differentiate between the fully integrated, internal model, and external entities which may assume the most different forms, ranging from a non-profit, participated organization to a consortial form of networked, interconnected TTOs \citep{Brescia2016}. These external entities are a particular useful solution in the case of smaller universities and fragmented environments \citep{Debackere2005}.

\subsection{Organizational models and objectives}

Even more important than the organizational structure, is the organizational role of TTOs: these offices must incentive researchers to engage in technology transfer activities, through their impact on the academic environment. Starting from \citet{Jensen1998}, many authors have modelled the TTO role as an agent of both the university administration, faculties and researchers, involved in balancing the needs of every actor. In this perspective, the orientation and the activity of the office both improve the outcome of the technology transfer process and the raw material the office has to work with, the invention disclosure \citep{Siegel2007}. 

More specifically, apart from the obvious impact of university-level regulation and incentives, TTOs have a peculiar, major factor at their disposal, to enhance the attitude of scientists to disclose: their reputation. By demonstrating a deep understanding of both academics and firms, and by building a history of successes, TTOs can make themselves perceived as professionalized and useful, thus acquire credibility and gaining the trust of academics \citep{OwenSmith2001}. Specific to the importance of the ability to understand firms and markets, \citet{Muscio2010} empirically demonstrate that a non-academic background of the TTO leaders is linked to a greater use of the TTOs services.

Otherwise, faculty and academics may prefer to circumvent the TTO by establishing a direct relationship with the industry or refusing to disclose inventions, thus leading to a waste in resources and a possible underperformance driven from the non-use of professionalized services. The same result may be reached if previous collaborations with the TTO have led to unsatisfactory outcomes, or if the TTO cannot offer a valuable service due to lack or inadequateness of personnel. 

The way TTOs relate with academics largely depend itself on the goals set by the university administration. Two main model are recognized by the literature: the revenue maximization and the diffusion maximization models. The first, the most common, will make TTOs focus on short-term cash maximization, where a great risk aversion may lead to \enquote{suboptimal licensing strategies} \citep{Markman2005}. Moreover, \citet{Siegel2003} identify a mismatch between this commercialization model and the motivations that drive faculty and academics involvement, thus a lower performance.

The second model, apart from being preferred by scientists, would be the natural path for technology transfer officers: a survey by \citet{Jensen1998}, demonstrate that the main objectives of technology managers are the number of inventions commercialized and of licenses executed. However, this model does not fit with the university's actual need for research funds from other sources than the government. This may be the main factor explaining why the technology manager interviewee see themselves as \enquote{juggling} between different interests, trying to maximize a weighted average of the administration's utility and the inventor's utility.

\citet{Fitzgerald2015} use a quite different approach to uncover drivers and models beneath TTOs activities: their mission statements, that are an instrument to signal to stakeholders the main, long term mission of the organizations. In concept, they should be useful also as a guide for decision making, as a tool for the formulation and implementation for a strategic plan, thus influencing the personnel behavior and their performance.

He reported in his empirical research that even if almost all mission statements report the identification of main products, services, customers and markets, almost no one reported the desired public image, self-concept, expression of commitment or any element of the organizational philosophy. These results may suggest that the impact of the organizational role of TTOs, especially as motivator for academics, is still underestimated or, worse, poorly understood. 

\subsection{Activities and personnel}

The office's objectives and model directly influence the type of activities it will perform, as well as the extent of the effort dedicated to each activity. In earlier days,\citet{Siegel2003a} recognized as main objectives licenses and royalties, from the patent activities, and the support to firms and scientists for economic and product development. Similarly, \citet{OwenSmith2001} reported that the resource constrain makes TTOs concentrate in the core activity, identified in the management of the IP portfolio.

More recent research reported a shift, or at least an enlargement, of the TTO main activity: \citet{Geuna2009} reported as main focuses (1) patent and licensing activities and (2) the creation of spinoff. Even if he mentioned also contract research and consultancy, it is important to note that in his view the focus has shifted from the evergreen, well-known patent affair, to more generic activities that imply a commercial exploitation of university knowledge directly performed by internal personnel. This perspective is reinforced by \citet{Balderi2010} who identified as main activities patenting and licensing, spinoff and other activities complementary to the first ones.

A recent and useful framework can be taken form \citet{Alexander2013}who linked TTO's core activities to four key objectives: 
\begin{itemize}
\item Facilitate the management of activities and projects: this require a continue assistance to academics and firms from the setup to the control of key stages and follow-up;
\item Enabling the transfer of intellectual property and facilitate entrepreneurial activities, i.e. the management of the patent process, administrative and bureaucratic support for spinoffs;
\item Promote and develop knowledge-based support services, ranging from contract and cooperative research to training and personal development;
\item Establish knowledge-based boundary-spanning activities, both in terms of stand-alone knowledge and through mobility and networking. 
\end{itemize}

A more debated topic refers to the disclosure eliciting activity. It is true that nowadays quite all academics engage technology transfer activities only through the TTO, letting the office controlling and managing the relation, but some researchers still not disclose inventions. It may be required by the university regulation or other external sources, i.e. the case of the Bayh-Dole act, but these rules are rarely enforced. On one hand, there is a clear difficult in monitoring research activities and comprehend which results may be susceptible of commercialization, while on the other hand TTOs usually have resources and time constrains that impede the eliciting activity. 

This scenario highlight the importance for TTOs of having highly skilled and trained professionals. Every activity requires specific knowledge and often interdisciplinary competences: example are IP and contract lawyers, business developers, business analysts, marketing personnel. They must comprehend both the academic and the business world, preferably with experience in both, as well as be able to evaluate technologies embedded in each proposal. Skilled personnel may gain a good reputation and enhance the disclosure, while dissatisfaction may lead academic to circumvent the TTO. 

To ensure a good performance, TTO may also consider the provision of specific services from external provider, both specialized in one activity, i.e. patent attorneys and technology consultant, as well as specialized in the technology transfer itself, i.e. TT centers, consortia etc. Another alternative is to develop dedicated external support facilities, as business incubators and science parks.

\section{Faculty}

Faculty and departments may have an important role in the technology transfer process, both as a group of academics that share a common research theme and as an organizational intermediary between the central administration and scientists. 

A first, relevant factor is the faculty leadership: as stated before, an experienced, supportive leader positively influence the academics perception of the university orientation toward commercial activities. \citet{Guerrero2014} reinforce this perspective by analyzing the importance of the attitudes of key actors, like faculty member and leaders. 

Similarly, \citet{Muscio2010} found significant the faculty leadership (identified in the department director) trust in the TTO, which increase the probability of a successful exploitation – assuming a positive impact of the office's professionalization; he also found an inverse correlation between the leader age, as proxy of experience and previous involvement with industry, and the probability of trusting and contacting the TTO. However, faculty leadership is expected to execute policies in compliance with the university mission and objectives \citep{Chang2016}, therefore linking the its supportiveness to general policies and the administration. 

A fundamental study on this topic is the one authored by \citet{OwenSmith2001}. He found three significant factors for the institutional success of patenting, as one of the technology transfer channels: (1) the faculty perception on the potential benefit of patenting; (2) the perception of time and resource required to interact with the TTO, as well as its perceived quality; (3) the general opinion of the university on the technology transfer topic. Along with perceptions, faculty behavior is influenced by the institutional environment and its organizational structure, thus influencing the individual behavior. 

Firstly, promotion and tenure policies may be shaped by the university mission and orientation, but are enforced by faculties and departments. Moreover, they are also in charge for the evaluation of research personnel, both on scientific outcomes and commercial exploitation \citep{Chang2016}. Secondly, social norms and practices established at the university level may strongly influence the faculty attitude; should be remembered that, according to \citet{Bercovitz2006}, a shift can be made by setting the correct incentives, especially at the faculty level, regardless the university history.

Another relevant faculty variable is its scale, both in term of human and technical resources; as an example, \citet{OwenSmith2001} empirical findings suggests that, ceteris paribus, best performers have \enquote{more researchers and more resources devoted to research}. Similarly, the size of the department is suggested to have a U-shaped relationship with the volume of industry interaction \citep{DEste2007}, where small departments may have not enough resources and large departments tend to engage a more basic research, far from industry problematics.

Past performance is found to be useful in predicting future ones. A useful proxy is the departmental research income, per member, received from industry which is found to be correlated with future engagement in commercialization activities \citep{DEste2007}. Also \citet{Blumenthal1996} in earlier days found evidences that industry-funded faculty members are commercially more productive, suggesting a Matthew effect. Similarly, \citet{Thursby2002} empirical findings indicate that a growth in faculty propensity to disclose inventions, itself one of the most significant predictors of technology transfer performance, is clearly linked to past licensing success. 

Of similar impact is the overall past scientific performance of the faculty \citep{OShea2005}. \citep{DEste2011} indicate that, apart from its effect on the volume of interaction, it influences also the preferred channel; specifically, he found that researchers in lower rated departments tend to prefer consulting, whereas departments with higher evaluations and income are more likely to engage in more frequent contract research.

The impact of the faculty's scientific field has been extensively studied. Firstly, there is no doubt that cultural norms across fields may be important \citep{DEste2007}; secondly, the nature of the research – its applicability in industry matters – is critical in determining the extent and volume of commercialization activities. In fact, some research fields easily meet specific market conditions, and \citet{OShea2005} found that faculties and departments involved in such research usually receive more funding, either from industry, government or the university administration.

Oppositely, faculties specialized in basic research may prefer to not disclose inventions, because afraid or unwilling to spend time and resources on the successive applied research needed for a commercial exploitation \citep{Bercovitz2006}. Basic research also refers more strictly to the open science paradigm, leading such faculties to avoid commercialization due to possible publication delays for patenting and marketing purposes. 

Inside the basic and applied research areas, the literature is more divided around the effect of specific disciplines on the commercial engagement; different authors provide contradictory findings on whether the research topic, as in biotech, medical school, engineering, increase technology transfer. However, \citet{OwenSmith2001} observe that in any case, the value of patents may vary significantly across areas, in terms of the protection extent, for leverage potential and as source of incomes. 

Another significant research on this topic refer to the impact of research areas and the cognitive distance from the related industry: surprisingly, \citet{Muscio2010} found that departments with greater cognitive distance from industrialists do collaborate more. This finding support the theoretical literature that suggested an inverted U-shaped relationship between cognitive distance and university-industry collaboration.

\citet{Wong2010} also found significant the impact of faculty internationalization on patenting performance, but results are ambiguous: in North America it has a negative effect, while positive elsewhere. These findings suggest further investigations on which faculty factors, among different university systems, can lead to this contradictory result. 

Finally, faculty benefits from technology transfer usually refer to the individual benefits, and the impact that these could have on the research performance of the overall personnel. \citet{Baldini2007}, for examples, indicate as major benefit the direct access to industry knowledge, laboratories and funds, the positive influence on researchers' career and earnings, new and different ways to exploit the researchers' abilities, factors previously seen as motivations for academics.

Again, also departments should demonstrate ambidexterity in order to gain a good overall performance. It is also fundamental to note that this ambidexterity is largely influenced by the department perception of the institutional flexibility, and may have a positive, significant influence on the individual ambidexterity. This setting helps clarify the great extent of environmental and institutional impact on the probability of engaging in technology transfer.
